---
title: "Gathering r4d data"
---

🚧 Attention: Work-in-progress. 🚧

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(here)
library(janitor)
library(yaml)
library(glue)
library(readxl)
p3_files <- read_yaml(here("inst", "extdata", "p3-tables.yml"))
download_table_from_p3 <- function(csv_table, dest_folder = "data-raw") {
  url_table <- glue("http://p3.snf.ch/P3Export/{csv_table}") %>%
    download.file(here(dest_folder, csv_table))
}
read_p3_file <- function(file_name, loc = "data-raw") {
  here(loc, file_name) %>% 
  read_delim(";", escape_double = FALSE, col_types = cols(.default = "c")) %>%
  clean_names()
}
```


## Read data from P3

Downloading raw data from [P3](http://p3.snf.ch/Pages/DataAndDocumentation.aspx) into [data-raw](./data-raw/).

```{r to-be-tested, eval=FALSE}
walk(p3_files, download_table_from_p3)
```

## Read local P3 files

```{r, eval=FALSE}
projects_raw <- read_p3_file(pluck(p3_files, 1))
people_raw <- read_p3_file(pluck(p3_files, 2))
```

## Logic

Finding 86 projects (77 [r4d.ch](http://www.r4d.ch/r4d-programme/all-projects) and 9 SPIRIT). 62 distinct project titles in [r4d.ch](http://www.r4d.ch/r4d-programme/all-projects) and 9 distinct project titles in SPIRIT.

```{r projects, eval=FALSE}
projects <- projects_raw %>% 
  filter(str_detect(funding_instrument, "r4d|SPIRIT")) %>% # MVP
  select(project_number, 
         project_title,
         funding_instrument,
         start_date,
         end_date,
         approved_amount,
         institution,
         university,
         keywords) %>%
  mutate(
    start_date = dmy(start_date),
    end_date = dmy(end_date),
    running = today() %within% interval(start_date, end_date),
    start_date = round_date(start_date, unit = "months"),
    end_date = round_date(end_date, unit = "months"),
    start_date = format(start_date, "%b %Y"),
    end_date = format(end_date, "%b %Y"),
    approved_amount = as.numeric(approved_amount),
    approved_amount = round(approved_amount),
    approved_amount = as.integer(approved_amount))

write_csv(projects, here("data", "projects.csv"))
```

```{r, disciplines, eval=FALSE}
projects <- read_csv(here("data", "projects.csv")) # p3 project data on r4d
disciplines <-  projects_raw %>%
  semi_join(mutate(projects, project_number = as.character(project_number)),
                   by = "project_number")  %>%
  select(project_number, discipline_name_hierarchy, discipline_name) %>%
  unite(col = "disciplines", starts_with("discipline_"), sep = ";") %>%
  mutate(disciplines = str_split(disciplines, ";")) %>%
  unnest(disciplines)

test_count <- disciplines %>% count(project_number)
if (all(pull(test_count, n) == 3)) {
  disciplines <- disciplines %>% mutate(tier = rep(1:3, nrow(test_count)))
  write_csv(disciplines, here("data", "disciplines.csv"))
}
```


Add detailed institute data missing from the CSV data dumps by scrapping P3.

```{r project-details, eval=FALSE}
scrape_p3_project <- function(project_number) {
  message(glue("* Scrapping {project_number}"))
  Sys.sleep(.25)
  
  p3_url <- glue("http://p3.snf.ch/project-{project_number}")
  html_page <- xml2::read_html(p3_url)
  project_people <- html_page %>%
    rvest::html_nodes(".person-field-0 a") %>%
    rvest::html_attr("href")
  
  project_people <- html_page %>%
    rvest::html_nodes(".person-field-0 a") %>%
    rvest::html_attr("href")
  project_institutes <- html_page %>%
    rvest::html_nodes(".person-field-1 a") %>%
    rvest::html_attr("href")

  tibble(
    people_p3slug = project_people,
    institute_p3slug = project_institutes) %>%
    extract(people_p3slug, 
            "person_id_p3", 
            regex = "(\\d+)", 
            remove=FALSE) %>%
    extract(institute_p3slug, 
            "institute_id_p3", 
            regex = "(\\d+)", 
            remove=FALSE) %>%
    mutate(project_id = project_number) %>%
    select(project_id, everything())
}

p3_project_details <- projects %>% 
  pull(project_number) %>% 
  map_df(scrape_p3_project)

write_csv(p3_project_details, here("data-raw", "p3-grant-details.csv"))
```

```{r institute-details, eval=FALSE}
scrape_p3_institute <- function(institute_number) {
  message(glue("* Scrapping /institute-{institute_number}"))
  Sys.sleep(.25)
  p3_url <- glue("http://p3.snf.ch/institute-{institute_number}")
  html_page <- xml2::read_html(p3_url)
  institute_details <- 
    html_page %>%
    rvest::html_nodes(".institute td") %>%
    rvest::html_text() %>%
    str_squish() %>%
    head(2)

  tibble(
    institute_p3_name = pluck(institute_details, 1),
    institute_p3_addr = pluck(institute_details, 2)
  ) %>%
    mutate(institute_id = institute_number) %>%
    select(institute_id, everything())
}

p3_institute_details <- p3_project_details %>%
  filter(institute_id_p3 != "0") %>%
  distinct(institute_id_p3) %>%
  pull(institute_id_p3) %>%
  map_df(scrape_p3_institute)

write_csv(p3_institute_details, here("data-raw", "p3-institute-details.csv"))
```

```{r people, eval=FALSE}
people <- people_raw %>%
  pivot_longer(
    starts_with("projects"), 
    names_to = "role", 
    values_to = "project_number",
    values_drop_na = TRUE) %>%
  mutate(project_number = str_split(project_number, ";")) %>%
  unnest(project_number) %>%
  semi_join(projects, "project_number") %>%
  select(project_number,
         person_id_snsf,
         role,
         institute_name,
         institute_place) %>%
  left_join(select(p3_project_details, 
                   project_id,
                   person_id_p3,
                   institute_id_p3), 
            by = c("project_number" = "project_id", 
                   "person_id_snsf" = "person_id_p3")) %>%
  mutate(
    role = str_remove(role, "projects_as_"),
    institute_id_p3 = na_if(institute_id_p3, "0"))

write_csv(people, here("data", "people.csv"))


```

```{r, eval=FALSE}
# for completeness
collab <- read_p3_file(pluck(p3_files, 3))

# 771 more collaborations...
more <- semi_join(collab, r4d, "project_number")
# more %>% count(group_person, sort = TRUE) %>% print(n = 50)
# more %>% filter(str_detect(group_person, "CO-APPLICANT"))
```


```{r load-p3-details, eval=FALSE}
p3_project_details <- read_csv(here("data-raw", "p3-grant-details.csv"))
p3_institute_details <- read_csv(here("data-raw", "p3-institute-details.csv"))

# tests :)
# people %>% anti_join(p3_project_details, by = c("person_id_snsf" = "person_id_p3"))
# p3_project_details %>% anti_join(people, by = c("person_id_p3" = "person_id_snsf"))
```


Preparing institute data for Google maps querying.

```{r gmaps-ready, eval=FALSE}
df_institutes <- people %>%
  filter(!is.na(institute_id_p3)) %>%
  select(institute_id_p3, institute_name, institute_place) %>%
  distinct(institute_id_p3, .keep_all = TRUE) %>%
  mutate(
    institute_name = str_remove_all(institute_name, "[[:punct:]]$"),
    institute_name = str_squish(institute_name),
    institute = glue("{institute_name}, {institute_place}"),
    institute = str_remove_all(institute, "\\d+"),
    institute = str_squish(institute),
    institute = str_remove_all(institute, "[[:punct:]]$"),
    institute = str_squish(institute),
    institute_ascii = stringi::stri_trans_general(institute, "Latin-ASCII"),
    institute_ascii = str_replace_all(institute_ascii, fixed(" -"), ","),
    institute_ascii = str_replace_all(institute_ascii, fixed(" ("), ", "),
    institute_ascii = str_replace_all(institute_ascii, fixed("/"), ", "),
    institute_ascii = str_replace_all(institute_ascii, fixed(")"), ","))

# Write 
write_csv(df_institutes, here("data-raw", "df_institutes.csv"))
```

## Query Google Maps 

```{r, eval=FALSE}
api_key <- Sys.getenv("GMAPS_API")
# https://developers.google.com/maps/documentation/places/web-service/overview
```

### Places Search

```{r, eval=FALSE}
gmaps <- tibble(
  gm_status = as.character(),
  gm_icon = as.character(),
  gm_lat = as.integer(),
  gm_lng = as.integer(),
  gm_plus = as.character(),
  gm_addr = as.character(),
  gm_name = as.character(),
  gm_id = as.character(),
  gm_types = as.character(),
  institute_id = as.character())

ntotal <- nrow(df_institutes)
df_institutes <- df_institutes %>%
  mutate(institute_urlenc = map_chr(institute_ascii, URLencode))

for(k in 11:nrow(df_institutes)) {
  Sys.sleep(.75) # 500ms
  message(glue("* Quering row {k} of {ntotal}.."))
  inst_id <- purrr::pluck(df_institutes, "institute_id_p3", k)
  url_enc <- purrr::pluck(df_institutes, "institute_urlenc", k)
  url_api <- glue("https://maps.googleapis.com/maps/api/",
                  "place/findplacefromtext/json?input={url_enc}&",
                  "inputtype=textquery&",
                  "fields=business_status,icon,plus_code,",
                  "formatted_address,geometry,name,place_id,types&",
                  "key={api_key}")
  # api call
  gm_res <- jsonlite::fromJSON(url_api)
  if (purrr::pluck(gm_res, "status") == "OK") {
    row_status <-  gm_res %>%
      purrr::pluck("candidates", "business_status", 1)
    row_icon <-  gm_res %>%
      purrr::pluck("candidates", "icon", 1)
    row_lat <-  gm_res %>%
      purrr::pluck("candidates", "geometry", "location", "lat", 1)
    row_lng <-  gm_res %>%
      purrr::pluck("candidates", "geometry", "location", "lng", 1)
    row_plus <-  gm_res %>%
      purrr::pluck("candidates", "plus_code", 1) %>%
      head(1)
    row_addr <-  gm_res %>%
      purrr::pluck("candidates", "formatted_address", 1)
    row_name <-  gm_res %>%
      purrr::pluck("candidates", "name", 1)
    row_id <- gm_res %>%
      purrr::pluck("candidates", "place_id", 1)
    row_types <-  gm_res %>%
      purrr::pluck("candidates", "types", 1) %>% 
      str_c(collapse = ",")
    gmaps <- gmaps %>%
      add_row(
        gm_status = if_else(is_null(row_status), NA_character_, row_status),
        gm_icon = if_else(is_null(row_icon), NA_character_, row_icon),
        gm_lat = if_else(is_null(row_lat), NA_real_, row_lat),
        gm_lng = if_else(is_null(row_lng), NA_real_, row_lng),
        gm_plus = if_else(is_null(row_plus), NA_character_, row_plus),
        gm_addr = if_else(is_null(row_addr), NA_character_, row_addr),
        gm_name = if_else(is_null(row_name), NA_character_, row_name),
        gm_id = if_else(is_null(row_id), NA_character_, row_id),
        gm_types = if_else(is_null(row_types), NA_character_, row_types),
        institute_id = inst_id)
  }
  else(
    message("!!!API KO")
  )
}

write_csv(gmaps, here("data-raw", "gmaps.csv"))
```


```{r, countries, eval=FALSE}
read_csv(here("data-raw", "gmaps.csv")) %>%
  select(
    institute_id,
    gmaps_id = gm_id,
    gmaps_lat = gm_lat,
    gmaps_lng = gm_lng,
    gmaps_name = gm_name,
    gmaps_addr = gm_addr) %>%
  # extract country from address
  mutate(
    gmaps_addr_ascii = stringi::stri_trans_general(
      gmaps_addr, 
      "Latin-ASCII")) %>%
  extract(
    gmaps_addr_ascii, 
    "gmaps_country", 
    regex = "(?<=[,] )([[:space:][:alpha:].-]+$)") %>%
  write_csv(here("data", "gmaps.csv"))
```


Add "difficult" cases separately (TODO).

```{r google-missing, eval=FALSE}
# TODO: failsafe for 11 missing values

df_institutes %>% 
  anti_join(gmaps, by = c("institute_id_p3" = "institute_id")) %>%
  pull(institute_ascii)
```

```{r google-details, eval=FALSE}
# Add websites
```

## DAC Country List

```{r}
# https://www.oecd.org/development/financing-sustainable-development/development-finance-standards/dacandcrscodelists.htm
# https://www.oecd.org/dac/financing-sustainable-development/development-finance-standards/DAC-CRS-CODES.xls

```

```{r, eval=FALSE}
file_name <- here("data-raw", 
                  "r4d_researchers_institutions_countries.xlsx")
sheets <- excel_sheets(path = file_name)
dfs <- sheets %>%
  map_df(function(x)
    readxl::read_excel(file_name, sheet = x) %>% select(1:6)) %>%
  clean_names()
# only 45 projects
```


```{r, eval=FALSE}
file_name <- here("about", "r4d.html")
html <- read_lines(file_name)
html <- str_subset(html, "DocId")
docid <- str_extract_all(html, '(?<=\\"DocId\\":)[[:digit:]]{4}') %>%
  flatten_chr()

data_list <- html %>% 
  str_split(',"') %>%
  map(~ str_replace_all(.x, '\\\\u002f', '/')) %>%
  map(~ str_remove_all(.x, '\\"')) %>%
  map(str_split_fixed, pattern = ":", n = 2) %>%
  map(~ .x[2:4, 2])

data_df <- tibble(
  doc_id = map_chr(data_list, ~ .x[1]),
  title = map_chr(data_list, ~ .x[2]),
  url = map_chr(data_list, ~ .x[3])
)


data_df <- data_df %>% 
  mutate(title_simple = str_to_lower(title))

grants <- grants %>% 
  mutate(title_simple = str_to_lower(project_title))

data_df %>% left_join(grants, "title_simple")

grants %>% filter(project_number == "160906") %>%
  pull(title_simple)

data_df %>% head(1) %>%
  pull(title_simple)

```

