---
title: "Gathering r4d data"
---

🚧 Attention: Work-in-progress. 🚧

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(here)
library(janitor)
library(yaml)
library(glue)
library(readxl)
p3_files <- read_yaml(here("inst", "extdata", "p3-tables.yml"))
# wrapper function to download the CSV files from the P3 website
download_table_from_p3 <- function(csv_table, dest_folder = "data-raw") {
  url_table <- glue("http://p3.snf.ch/P3Export/{csv_table}") %>%
    download.file(here(dest_folder, csv_table))
}
# specific for P3 CSV files
read_p3_file <- function(file_name, loc = "data-raw") {
  here(loc, file_name) %>%
    read_delim(";", escape_double = FALSE, col_types = cols(.default = "c")) %>%
    clean_names()
}
```


## Read data from P3

Downloading raw data from [P3](http://p3.snf.ch/Pages/DataAndDocumentation.aspx) into [data-raw](./data-raw/).

```{r to-be-tested, eval=FALSE}
walk(p3_files, download_table_from_p3)
```

## Read local P3 files

```{r, eval=FALSE}
projects_raw <- read_p3_file(pluck(p3_files, 1))
people_raw <- read_p3_file(pluck(p3_files, 2))
```

## Logic

Finding 92 projects (77 [r4d.ch](http://www.r4d.ch/r4d-programme/all-projects) and 15 SPIRIT). 62 distinct project titles in [r4d.ch](http://www.r4d.ch/r4d-programme/all-projects) and 15 distinct project titles in SPIRIT.

```{r projects, eval=FALSE}
projects <- projects_raw %>%
  filter(str_detect(funding_instrument, "r4d|SPIRIT")) %>%
  select(
    project_number,
    project_title,
    funding_instrument,
    start_date,
    end_date,
    approved_amount,
    institution,
    university,
    keywords
  ) %>%
  mutate(
    start_date = dmy(start_date),
    end_date = dmy(end_date),
    running = today() %within% interval(start_date, end_date),
    start_date = round_date(start_date, unit = "months"),
    end_date = round_date(end_date, unit = "months"),
    start_date = format(start_date, "%b %Y"),
    end_date = format(end_date, "%b %Y"),
    approved_amount = as.numeric(approved_amount),
    approved_amount = round(approved_amount),
    approved_amount = as.integer(approved_amount)
  )

write_csv(projects, here("data", "projects.csv"))
```

### Create tags table for research earth

```{r tags, eval=FALSE}
projects_tags <-
  projects %>%
  select(
    project__id = project_number,
    tags = keywords
  ) %>%
  arrange(project__id) %>%
  mutate(tags = str_split(tags, "; ")) %>%
  unnest(tags) %>%
  mutate(tags = str_split(tags, ", ")) %>%
  unnest(tags)

tags <-
  projects_tags %>%
  mutate(tags_low = str_to_lower(tags)) %>%
  distinct(tags_low, .keep_all = TRUE) %>%
  mutate(tags_name = if_else(tags == tags_low,
    str_to_title(tags),
    tags
  )) %>%
  mutate(tags_id = row_number()) %>%
  select(-project__id)

tags %>%
  select(tags_id, tags_name) %>%
  write_csv(here("data", "research_earth", "tags.csv"))

# lookup table for linking projects to the tags table
projects_tag_id <-
  projects_tags %>%
  left_join(tags, by = c("tags")) %>%
  group_by(project__id) %>%
  summarize(tag_id = str_c(tags_id, collapse = ","))
```

### Create disciplines table

```{r, disciplines, eval=FALSE}
projects <- read_csv(here("data", "projects.csv")) # p3 project data on r4d
disciplines <- projects_raw %>%
  semi_join(mutate(projects, project_number = as.character(project_number)),
    by = "project_number"
  ) %>%
  select(project_number, discipline_name_hierarchy, discipline_name) %>%
  unite(col = "disciplines", starts_with("discipline_"), sep = ";") %>%
  mutate(disciplines = str_split(disciplines, ";")) %>%
  unnest(disciplines)

test_count <- disciplines %>% count(project_number)
if (all(pull(test_count, n) == 3)) {
  disciplines <- disciplines %>% mutate(tier = rep(1:3, nrow(test_count)))
  write_csv(disciplines, here("data", "disciplines.csv"))
}
```

### Create disciplines table for research earth

```{r, disciplines-re, eval=FALSE}
disciplines_names_numbers <-
  projects_raw %>%
  distinct(discipline_name, discipline_number) %>%
  arrange(discipline_name)

ref_disciplines <-
  disciplines %>%
  distinct(disciplines, tier) %>%
  rename(discipline_name = disciplines) %>%
  left_join(disciplines_names_numbers, by = "discipline_name") %>%
  # clean legacy name inconsistencies in the P3 data
  mutate(discipline_clean = if_else(is.na(discipline_number),
    str_extract(
      discipline_name,
      "[[:alpha:]]+(?=,)"
    ),
    discipline_name
  )) %>%
  select(discipline_name, discipline_clean, tier) %>%
  left_join(
    disciplines_names_numbers,
    by = c("discipline_clean" = "discipline_name")
  ) %>%
  arrange(discipline_number, tier)

ref_disciplines %>%
  select(
    disciplines_id = discipline_number,
    disciplines_name = discipline_clean,
    tier
  ) %>%
  distinct(disciplines_id, .keep_all = TRUE) %>%
  write_csv(here("data", "research_earth", "disciplines.csv"))

# lookup table for linking projects to the disciplines table
projects_disciplines_id <-
  disciplines %>%
  select(
    project__id = project_number,
    discipline_name = disciplines
  ) %>%
  left_join(ref_disciplines, by = "discipline_name") %>%
  distinct(project__id, discipline_number) %>%
  group_by(project__id) %>%
  summarize(disciplines_id = str_c(discipline_number, collapse = ","))
```



Add detailed institute data missing from the CSV data dumps by scrapping P3.

```{r project-details, eval=FALSE}
scrape_p3_project <- function(project_number) {
  message(glue("* Scrapping {project_number}"))
  Sys.sleep(.25)

  p3_url <- glue("http://p3.snf.ch/project-{project_number}")
  html_page <- xml2::read_html(p3_url)
  project_people <- html_page %>%
    rvest::html_nodes(".person-field-0 a") %>%
    rvest::html_attr("href")

  project_people <- html_page %>%
    rvest::html_nodes(".person-field-0 a") %>%
    rvest::html_attr("href")
  project_institutes <- html_page %>%
    rvest::html_nodes(".person-field-1 a") %>%
    rvest::html_attr("href")

  tibble(
    people_p3slug = project_people,
    institute_p3slug = project_institutes
  ) %>%
    extract(people_p3slug,
      "person_id_p3",
      regex = "(\\d+)",
      remove = FALSE
    ) %>%
    extract(institute_p3slug,
      "institute_id_p3",
      regex = "(\\d+)",
      remove = FALSE
    ) %>%
    mutate(project_id = project_number) %>%
    select(project_id, everything())
}

p3_project_details <- projects %>%
  pull(project_number) %>%
  map_df(scrape_p3_project)

write_csv(p3_project_details, here("data-raw", "p3-grant-details.csv"))
p3_project_details <- read_csv(here("data-raw", "p3-grant-details.csv"))

# lookup table for linking projects to the organizations table
projects_organizations_id <-
  p3_project_details %>%
  select(
    project__id = project_id,
    institute_id_p3
  ) %>%
  mutate(project__id = as.character(project__id)) %>%
  filter(institute_id_p3 != "0") %>%
  group_by(project__id) %>%
  summarize(organizations_id = str_c(institute_id_p3, collapse = ","))
```

```{r institute-details, eval=FALSE}
scrape_p3_institute <- function(institute_number) {
  message(glue("* Scrapping /institute-{institute_number}"))
  Sys.sleep(.25)
  p3_url <- glue("http://p3.snf.ch/institute-{institute_number}")
  html_page <- xml2::read_html(p3_url)
  institute_details <-
    html_page %>%
    rvest::html_nodes(".institute td") %>%
    rvest::html_text() %>%
    str_squish() %>%
    head(2)

  tibble(
    institute_p3_name = pluck(institute_details, 1),
    institute_p3_addr = pluck(institute_details, 2)
  ) %>%
    mutate(institute_id = institute_number) %>%
    select(institute_id, everything())
}

p3_institute_details <- p3_project_details %>%
  filter(institute_id_p3 != "0") %>%
  distinct(institute_id_p3) %>%
  pull(institute_id_p3) %>%
  map_df(scrape_p3_institute)

write_csv(p3_institute_details, here("data-raw", "p3-institute-details.csv"))
```

```{r people, eval=FALSE}
people <- people_raw %>%
  pivot_longer(
    starts_with("projects"),
    names_to = "role",
    values_to = "project_number",
    values_drop_na = TRUE
  ) %>%
  mutate(project_number = str_split(project_number, ";")) %>%
  unnest(project_number) %>%
  mutate(project_number = as.numeric(project_number)) %>%
  semi_join(projects, "project_number") %>%
  select(
    project_number,
    person_id_snsf,
    role,
    institute_name,
    institute_place
  ) %>%
  mutate(person_id_snsf = as.numeric(person_id_snsf)) %>%
  left_join(select(
    p3_project_details,
    project_id,
    person_id_p3,
    institute_id_p3
  ),
  by = c(
    "project_number" = "project_id",
    "person_id_snsf" = "person_id_p3"
  )
  ) %>%
  mutate(
    role = str_remove(role, "projects_as_"),
    institute_id_p3 = na_if(institute_id_p3, "0")
  )

write_csv(people, here("data", "people.csv"))
```

```{r, eval=FALSE}
# for completeness
collab <- read_p3_file(pluck(p3_files, 3))

# 771 more collaborations...
more <- semi_join(collab, r4d, "project_number")
# more %>%count(group_person, sort = TRUE) %>%print(n = 50)
# more %>%filter(str_detect(group_person, "CO-APPLICANT"))
```


```{r load-p3-details, eval=FALSE}
p3_project_details <- read_csv(here("data-raw", "p3-grant-details.csv"))
p3_institute_details <- read_csv(here("data-raw", "p3-institute-details.csv"))

# tests :)
# people %>%anti_join(p3_project_details, by = c("person_id_snsf" = "person_id_p3"))
# p3_project_details %>%anti_join(people, by = c("person_id_p3" = "person_id_snsf"))
```


### Create projects table for research earth
```{r, eval=FALSE}
transdisciplinarity <-
  projects_raw %>%
  select(
    project__id = project_number,
    transdisciplinary = all_disciplines
  ) %>%
  mutate(project__id = as.numeric(project__id)) %>%
  semi_join(projects, by = c("project__id" = "project_number")) %>%
  mutate(
    project__id = as.character(project__id),
    transdisciplinary = str_detect(transdisciplinary, "/"),
    transdisciplinary = as.character(transdisciplinary),
    transdisciplinary = str_to_lower(transdisciplinary)
  )

projects_re <-
  projects %>%
  select(
    project__id = project_number,
    project_name = project_title,
    project_description = funding_instrument,
    start_date,
    end_date,
    approved_amount
  ) %>%
  mutate(
    project__id = as.character(project__id),
    email = "",
    projects_funder_id = "",
    start_date = str_replace(start_date, " 20", "-"),
    end_date = str_replace(end_date, " 20", "-")
  ) %>%
  left_join(projects_disciplines_id, by = "project__id") %>%
  left_join(projects_tag_id, by = "project__id") %>%
  left_join(
    select(
      projects_raw,
      project__id = project_number,
      contact_person = responsible_applicant
    ),
    by = "project__id"
  ) %>%
  left_join(transdisciplinarity, by = "project__id") %>%
  left_join(projects_organizations_id, by = "project__id")

projects_re %>%
  select(
    project__id,
    project_name,
    contact_person,
    email,
    start_date,
    end_date,
    approved_amount,
    project_description,
    transdisciplinary,
    disciplines_id,
    tag_id,
    projects_funder_id,
    projects_organizations_id = organizations_id
  ) %>%
  write_csv(here("data", "research_earth", "projects.csv"))
```

### Create funders table for research earth

```{r, eval=FALSE}
people %>%
  select(person_id_snsf, role) %>%
  distinct(person_id_snsf, .keep_all = TRUE) %>%
  mutate(person_id_snsf = as.character(person_id_snsf)) %>%
  left_join(select(people_raw, person_id_snsf, last_name, first_name),
    by = "person_id_snsf"
  ) %>%
  unite(funders_name, c("last_name", "first_name"), sep = " ") %>%
  mutate(
    funder_id = 1,
    funder_id = as.character(funder_id)
  ) %>%
  select(funder_id, funders_name, snsf = person_id_snsf, role) %>%
  write_csv(here("data", "research_earth", "funders.csv"))
```


Preparing institute data for Google maps querying.

```{r gmaps-ready, eval=FALSE}
df_institutes <- people %>%
  filter(!is.na(institute_id_p3)) %>%
  select(institute_id_p3, institute_name, institute_place) %>%
  distinct(institute_id_p3, .keep_all = TRUE) %>%
  mutate(
    institute_name = str_remove_all(institute_name, "[[:punct:]]$"),
    institute_name = str_squish(institute_name),
    institute = glue("{institute_name}, {institute_place}"),
    institute = str_remove_all(institute, "\\d+"),
    institute = str_squish(institute),
    institute = str_remove_all(institute, "[[:punct:]]$"),
    institute = str_squish(institute),
    institute_ascii = stringi::stri_trans_general(institute, "Latin-ASCII"),
    institute_ascii = str_replace_all(institute_ascii, fixed(" -"), ","),
    institute_ascii = str_replace_all(institute_ascii, fixed(" ("), ", "),
    institute_ascii = str_replace_all(institute_ascii, fixed("/"), ", "),
    institute_ascii = str_replace_all(institute_ascii, fixed(")"), ",")
  )

# Write
write_csv(df_institutes, here("data-raw", "df_institutes.csv"))
```

## Query Google Maps 

```{r, eval=FALSE}
api_key <- Sys.getenv("GMAPS_API")
# https://developers.google.com/maps/documentation/places/web-service/overview
```

### Places Search

```{r, eval=FALSE}
gmaps <- tibble(
  gm_status = as.character(),
  gm_icon = as.character(),
  gm_lat = as.integer(),
  gm_lng = as.integer(),
  gm_plus = as.character(),
  gm_addr = as.character(),
  gm_name = as.character(),
  gm_id = as.character(),
  gm_types = as.character(),
  institute_id = as.character()
)

ntotal <- nrow(df_institutes)
df_institutes <- df_institutes %>%
  mutate(institute_urlenc = map_chr(institute_ascii, URLencode))

for (k in 1:nrow(df_institutes)) {
  Sys.sleep(.75) # 500ms
  message(glue("* Quering row {k} of {ntotal}.."))
  inst_id <- purrr::pluck(df_institutes, "institute_id_p3", k)
  url_enc <- purrr::pluck(df_institutes, "institute_urlenc", k)
  url_api <- glue(
    "https://maps.googleapis.com/maps/api/",
    "place/findplacefromtext/json?input={url_enc}&",
    "inputtype=textquery&",
    "fields=business_status,icon,plus_code,",
    "formatted_address,geometry,name,place_id,types&",
    "key={api_key}"
  )
  # api call
  gm_res <- jsonlite::fromJSON(url_api)
  if (purrr::pluck(gm_res, "status") == "OK") {
    row_status <- gm_res %>%
      purrr::pluck("candidates", "business_status", 1)
    row_icon <- gm_res %>%
      purrr::pluck("candidates", "icon", 1)
    row_lat <- gm_res %>%
      purrr::pluck("candidates", "geometry", "location", "lat", 1)
    row_lng <- gm_res %>%
      purrr::pluck("candidates", "geometry", "location", "lng", 1)
    row_plus <- gm_res %>%
      purrr::pluck("candidates", "plus_code", 1) %>%
      head(1)
    row_addr <- gm_res %>%
      purrr::pluck("candidates", "formatted_address", 1)
    row_name <- gm_res %>%
      purrr::pluck("candidates", "name", 1)
    row_id <- gm_res %>%
      purrr::pluck("candidates", "place_id", 1)
    row_types <- gm_res %>%
      purrr::pluck("candidates", "types", 1) %>%
      str_c(collapse = ",")
    gmaps <- gmaps %>%
      add_row(
        gm_status = if_else(is_null(row_status), NA_character_, row_status),
        gm_icon = if_else(is_null(row_icon), NA_character_, row_icon),
        gm_lat = if_else(is_null(row_lat), NA_real_, row_lat),
        gm_lng = if_else(is_null(row_lng), NA_real_, row_lng),
        gm_plus = if_else(is_null(row_plus), NA_character_, row_plus),
        gm_addr = if_else(is_null(row_addr), NA_character_, row_addr),
        gm_name = if_else(is_null(row_name), NA_character_, row_name),
        gm_id = if_else(is_null(row_id), NA_character_, row_id),
        gm_types = if_else(is_null(row_types), NA_character_, row_types),
        institute_id = inst_id
      )
  } else {
    (
      message("!!!API KO")
    )
  }
}

write_csv(gmaps, here("data-raw", "gmaps.csv"))
```


```{r, countries, eval=FALSE}
read_csv(here("data-raw", "gmaps.csv")) %>%
  select(
    institute_id,
    gmaps_id = gm_id,
    gmaps_lat = gm_lat,
    gmaps_lng = gm_lng,
    gmaps_name = gm_name,
    gmaps_addr = gm_addr
  ) %>%
  # extract country from address
  mutate(
    gmaps_addr_ascii = stringi::stri_trans_general(
      gmaps_addr,
      "Latin-ASCII"
    )
  ) %>%
  extract(
    gmaps_addr_ascii,
    "gmaps_country",
    regex = "(?<=[,] )([[:space:][:alpha:].-]+$)"
  ) %>%
  write_csv(here("data", "gmaps.csv"))
```


### Create organizations table for research earth

```{r, eval=FALSE}
gmaps <- read_csv(here("data", "gmaps.csv"))

# look for organizations and projects
lookup_organizations_project_id <- p3_project_details %>%
  distinct(project_id, institute_id_p3) %>%
  filter(institute_id_p3 != 0) %>%
  group_by(institute_id_p3) %>%
  summarize(organizations_project_id = str_c(project_id, collapse = ",")) %>%
  mutate(institute_id_p3 = as.character(institute_id_p3)) %>%
  rename(organizations_id = institute_id_p3)

# lookup for organizations and disciplines

lookup_organizations_disciplines_id <- p3_project_details %>%
  distinct(project_id, institute_id_p3) %>%
  filter(institute_id_p3 != 0) %>%
  mutate(project_id = as.character(project_id)) %>%
  left_join(projects_disciplines_id, by = c("project_id" = "project__id")) %>%
  select(-project_id) %>%
  mutate(
    institute_id_p3 = as.character(institute_id_p3),
    disciplines_id = str_split(disciplines_id, ",")
  ) %>%
  unnest(disciplines_id) %>%
  distinct(institute_id_p3, disciplines_id) %>%
  group_by(institute_id_p3) %>%
  summarize(disciplines_id = str_c(disciplines_id, collapse = ",")) %>%
  rename(organizations_id = institute_id_p3)

# lookup for partner organisations

partners <- p3_project_details %>%
  distinct(project_id, institute_id_p3) %>%
  filter(institute_id_p3 != 0)

lookup_partners <- partners %>%
  full_join(partners, by = "project_id") %>%
  select(-project_id) %>%
  filter(institute_id_p3.x != institute_id_p3.y) %>%
  distinct(institute_id_p3.x, institute_id_p3.y) %>%
  group_by(institute_id_p3.x) %>%
  summarize(partner_organizations_id = str_c(institute_id_p3.y, collapse = ",")) %>%
  mutate(institute_id_p3.x = as.character(institute_id_p3.x)) %>%
  rename(organizations_id = institute_id_p3.x)

organizations <- gmaps %>%
  select(
    organizations_id = institute_id,
    organizations_name = gmaps_name,
    lat = gmaps_lat,
    lng = gmaps_lng,
    gmaps_id,
    address = gmaps_addr,
    country = gmaps_country,
  ) %>%
  mutate(
    organizations_id = as.character(organizations_id),
    website = "http://research-earth.de",
    organizations_funder_id = "1"
  ) %>%
  left_join(lookup_organizations_disciplines_id, by = "organizations_id") %>%
  left_join(lookup_partners, by = "organizations_id") %>%
  left_join(lookup_organizations_project_id, by = "organizations_id")

organizations %>%
  select(
    organizations_id,
    organizations_name,
    lat,
    lng,
    gmaps_id,
    address,
    country,
    website,
    organizations_disciplines_id = disciplines_id,
    organizations_funder_id,
    partner_organizations_id,
    organizations_project_id
  ) %>%
  write_csv(here("data", "research_earth", "organizations.csv"))
```

---

Add "difficult" cases separately (TODO).

```{r google-missing, eval=FALSE}
# TODO: failsafe for 11 missing values

df_institutes %>%
  anti_join(gmaps, by = c("institute_id_p3" = "institute_id")) %>%
  pull(institute_ascii)
```

```{r google-details, eval=FALSE}
# Add websites
```

## DAC Country List

```{r}
# https://www.oecd.org/development/financing-sustainable-development/development-finance-standards/dacandcrscodelists.htm
# https://www.oecd.org/dac/financing-sustainable-development/development-finance-standards/DAC-CRS-CODES.xls
```

```{r, eval=FALSE}
file_name <- here(
  "data-raw",
  "r4d_researchers_institutions_countries.xlsx"
)
sheets <- excel_sheets(path = file_name)
dfs <- sheets %>%
  map_df(function(x) {
    readxl::read_excel(file_name, sheet = x) %>% select(1:6)
  }) %>%
  clean_names()
# only 45 projects
```


```{r, eval=FALSE}
file_name <- here("about", "r4d.html")
html <- read_lines(file_name)
html <- str_subset(html, "DocId")
docid <- str_extract_all(html, '(?<=\\"DocId\\":)[[:digit:]]{4}') %>%
  flatten_chr()

data_list <- html %>%
  str_split(',"') %>%
  map(~ str_replace_all(.x, "\\\\u002f", "/")) %>%
  map(~ str_remove_all(.x, '\\"')) %>%
  map(str_split_fixed, pattern = ":", n = 2) %>%
  map(~ .x[2:4, 2])

data_df <- tibble(
  doc_id = map_chr(data_list, ~ .x[1]),
  title = map_chr(data_list, ~ .x[2]),
  url = map_chr(data_list, ~ .x[3])
)


data_df <- data_df %>%
  mutate(title_simple = str_to_lower(title))

grants <- grants %>%
  mutate(title_simple = str_to_lower(project_title))

data_df %>% left_join(grants, "title_simple")

grants %>%
  filter(project_number == "160906") %>%
  pull(title_simple)

data_df %>%
  head(1) %>%
  pull(title_simple)
```

